Seasonality in dominant upper-ocean dynamics at submesoscales

### 2016-05-08 writeup/rocha_etal.tex
Updates section 2 and 3

\section{The LLC2160 model output}
We use the output of the latitude-longitude polar cap (LLC)
MITgcm  numerical simulation. The output analyzed here, the LLC2160,
is the 1/24$^\circ$ --- 90-vertical-levels simulation from a hierarchy of global
forward numerical solutions on a LLC grid \citep{forget_etal2015} that were
spun up from a ECCO2 adjoint-method state estimate. The LLC2160 was forced by
tides and high-resolution surface fluxes similarly to the LLC4320 simulation
used by \citet{rocha_etal2016} --- see their appendix D. We chose the second-higher
resolution  simulation (LLC2160) of the LLC hierarchy because it was run for longer:
the LLC2160 output spans two years form March 2011 to April 2013. The model time-step
was 45 s and the model variables were saved hourly. LLC hierarchy model setup
and control files are available online
(at \texttt{http://mitgcm.org/viewvc/MITgcm/MITgcm$\_$contrib/
llc$\_$hires}). The model output was provided by D. Menemenlis (JPL/NASA).

A key aspects: (1) tides. \citep{sasaki_etal2014}.

\section{Results}

\subsection{The seasonal cycle in bulk quantities}

>>>>

\section{The LLC2160 model output}
We use the output of the latitude-longitude polar cap (LLC)
MITgcm  numerical simulation. The output analyzed here, the LLC2160,
is the 1/24$^\circ$ --- 90-vertical-levels simulation from a hierarchy of global
forward numerical solutions on a LLC grid \citep{forget_etal2015} that were
spun up from a ECCO2 adjoint-method state estimate. The LLC2160 was forced by
tides and high-resolution surface fluxes similarly to the LLC4320 simulation
used by \citet{rocha_etal2016} --- see their appendix D. We chose the second-higher
resolution  simulation (LLC2160) of the LLC hierarchy because it was run for longer:
the LLC2160 output spans two years form March 2011 to April 2013. The model time-step
was 45s and the model variables were saved hourly. LLC hierarchy setup
and control files are available online
(at \texttt{http://mitgcm.org/viewvc/MITgcm/MITgcm$\_$contrib/
llc$\_$hires}). For the LLC hierarchy, the sea-surface height is calculated
using a linearized equation. The model output was provided by D. Menemenlis
(JPL/NASA).

A key aspect of the simulation analyzed here is that it is forced by
the 16 most significant tidal components.
Barotropic tides interact with topography and generate internal
tides that project on mesoscales to submesoscales
\citep{rocha_etal2016}. Hence, tidal forcing fundamentally differentiates our analysis
from other modeling studies that investigated upper dynamics
and its seasonality \citep{sasaki_etal2014,qiu_etal2014}.

The region of focus is the northwest Pacific, in the vicinities of the Kuroshio
Extension, where previous studies have
shown strong seasonality in the quasi-gesotrophic surface velocity field.
We analyze a sub-domain of the LLC2160 simulation of about 2000 km$^2$
spanning 155-175$^\circ$E; 25-40$^\circ$N. We focus on the surface velocity field
and near-surface stratification.

\section{Seasonality in surface velocity and near-surface stratification}

We calculate the vertical vorticity, horizontal divergence, and the magnitude
of the lateral rate of strain:
\begin{equation}\label{vort}
\zeta \equiv v_x - u_y\, ,
\end{equation}
\begin{equation}\label{div}
\delta \equiv u_x + v_y\, ,
\end{equation}
and
\begin{equation}\label{strain}
S \equiv [(v_x+u_y)^2 + (u_x-v_y)^2]^{1/2}\, ,
\end{equation}
